---
title: "Spatial analysis TERVA2"
author: "Lea Mikkola"
date: "2023-12-04"
output:
  html_document:
    keep_md: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Remember to change the paths to your own directory after downloading the data.
```{r}
#install.packages('BiocManager')
#BiocManager::install('glmGamPoi')
setwd("~/Desktop/TERVA/TERVA2DATA/SPATIAL 2022/")
library(Matrix)
library(tidyverse)
library(Seurat)
library(patchwork)
library(hdf5r)
library(viridis)
library(clustree)
library(glmGamPoi)
library(gridExtra)

```

First load the data in with the Load_spatial R script or load an environment that contains the data.

```{r Loading-data}
# Load all Visium sections ####

#### AORTA ####

#A006D1
A006D1<- Load10X_Spatial( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/AO_D1/filtered_feature_bc_matrix/', 
                          filename = "filtered_feature_bc_matrix.h5", #name of the feature barcode matrix  
                          assay = "Spatial", 
                          slice = "V11L12098D1", # name of the tissue slice
                          filter.matrix = TRUE, # use only spots that under a tissue 
                          to.upper = FALSE, #converts all feature names to upper case
                          image = Read10X_Image( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/AO_D1/spatial/',  
                                                 image.name = "tissue_lowres_image.png", # name of the image file 
                                                 filter.matrix = TRUE, # use only spots under the tissue 
                                                 
                          ),
)

#A006C1
A006C1 <- Load10X_Spatial( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/AO_C1/filtered_feature_bc_matrix/',
                           filename = "filtered_feature_bc_matrix.h5", 
                           assay = "Spatial", 
                           slice = "V11L12098C1", 
                           filter.matrix = TRUE, 
                           to.upper = FALSE, 
                           image = Read10X_Image( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/AO_C1/spatial/',  
                                                  image.name = "tissue_lowres_image.png", 
                                                  filter.matrix = TRUE, 
                                                  
                           ),
)

#A004B1
A004B1 <- Load10X_Spatial( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/AO_B1/filtered_feature_bc_matrix/', 
                           filename = "filtered_feature_bc_matrix.h5", 
                           assay = "Spatial", 
                           slice = "V11L12098B1", 
                           filter.matrix = TRUE, 
                           to.upper = FALSE, 
                           image = Read10X_Image( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/AO_B1/spatial/',  
                                                  image.name = "tissue_lowres_image.png", 
                                                  filter.matrix = TRUE, 
                                                  
                           ),
)

#A004A1
A004A1 <- Load10X_Spatial( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/AO_A1/filtered_feature_bc_matrix/', 
                           filename = "filtered_feature_bc_matrix.h5", 
                           assay = "Spatial", 
                           slice = "V11L12098A1",
                           filter.matrix = TRUE, 
                           to.upper = FALSE, 
                           image = Read10X_Image( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/AO_A1/spatial/', 
                                                  image.name = "tissue_lowres_image.png", 
                                                  filter.matrix = TRUE, 
                                                  
                           ),
)

####CAROTID ARTERY####
#C006D1
C006D1<- Load10X_Spatial( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/CA_D1/filtered_feature_bc_matrix/', 
                          filename = "filtered_feature_bc_matrix.h5", #name of the feature barcode matrix  
                          assay = "Spatial", 
                          slice = "V11L12070D1", # name of the tissue slice
                          filter.matrix = TRUE, # use only spots that under a tissue 
                          to.upper = FALSE, #converts all feature names to upper case
                          image = Read10X_Image( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/CA_D1/spatial/',  
                                                 image.name = "tissue_lowres_image.png", # name of the image file 
                                                 filter.matrix = TRUE, # use only spots under the tissue 
                                                 
                          ),
)

#C006C1
C006C1 <- Load10X_Spatial( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/CA_C1/filtered_feature_bc_matrix/',
                           filename = "filtered_feature_bc_matrix.h5", 
                           assay = "Spatial", 
                           slice = "V11L12070C1", 
                           filter.matrix = TRUE, 
                           to.upper = FALSE, 
                           image = Read10X_Image( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/CA_C1/spatial/',  
                                                  image.name = "tissue_lowres_image.png", 
                                                  filter.matrix = TRUE, 
                                                  
                           ),
)

#C004B1
C004B1 <- Load10X_Spatial( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/CA_B1/filtered_feature_bc_matrix/', 
                           filename = "filtered_feature_bc_matrix.h5", 
                           assay = "Spatial", 
                           slice = "V11L12070B1", 
                           filter.matrix = TRUE, 
                           to.upper = FALSE, 
                           image = Read10X_Image( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/CA_B1/spatial/',  
                                                  image.name = "tissue_lowres_image.png", 
                                                  filter.matrix = TRUE, 
                                                  
                           ),
)

#C004A1
C004A1 <- Load10X_Spatial( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/CA_A1/filtered_feature_bc_matrix/', 
                           filename = "filtered_feature_bc_matrix.h5", 
                           assay = "Spatial", 
                           slice = "V11L12070BA1",
                           filter.matrix = TRUE, 
                           to.upper = FALSE, 
                           image = Read10X_Image( 'Visium TERVA/DATA Space Ranger outs H5 AND IMAGES/CA_A1/spatial/', 
                                                  image.name = "tissue_lowres_image.png", 
                                                  filter.matrix = TRUE, 
                                                  
                           ),
)

```


Create sample id's and then define a function to determine the names for plots later on. This requires that you define the names to your list of objects X with names(X) <- paste0(Y) where Y is list of names. 
```{r Sample-ID-and-functions, results='hide'}
A004A1@meta.data$sample_id <- "AOA1"
A004B1@meta.data$sample_id <- "AOB1"
A006C1@meta.data$sample_id <- "AOC1"
A006D1@meta.data$sample_id <- "AOD1" 

C004A1@meta.data$sample_id <- "CAA1"
C004B1@meta.data$sample_id <- "CAB1"
C006C1@meta.data$sample_id <- "CAC1"
C006D1@meta.data$sample_id <- "CAD1" 

getObjName <- function(x) {
  sampleID = x@meta.data[["sample_id"]][1]
  name = paste(sampleID)
  return(name)
}

makeX <- function(x){
  X <- c(A004A1, A004B1, A006C1, A006D1, C004A1, C004B1, C006C1, C006D1)
  Y <- c("A004A1", "A004B1", "A006C1", "A006D1", "C004A1", "C004B1", "C006C1", "C006D1") 
  names(X) <- paste0(Y)
  return(X)
}

savedir = getwd()


```


*QC*

```{r QC-Number-of-features, results='hide', fig.dpi=300, fig.width=50, fig.height=10}
X <- makeX()

L <- lapply(X, function(x) {
  x@meta.data[["percent_mito"]] <- PercentageFeatureSet(x, "^MT-")
  x@meta.data[["percent_ribo"]] <- PercentageFeatureSet(x, "^RP[SL]")
  return(x)
})

Y <- c("A004A1", "A004B1", "A006C1", "A006D1", "C004A1", "C004B1", "C006C1", "C006D1") 
names(L) <- paste0(Y)
list2env(L, envir = globalenv())
rm(L)

#A004A1.qcF <- A004A1[, A004A1$nFeature_Spatial > 40] #Test different thresholds and visually inspect what would be suitable level.
#A004B1.qcF <- A004B1[, A004B1$nFeature_Spatial > 40]
#A006C1.qcF <- A006C1[, A006C1$nFeature_Spatial > 90]
#A006D1.qcF <- A006D1[, A006D1$nFeature_Spatial > 90]
#004A1.qcF <- C004A1[, C004A1$nFeature_Spatial > 60] 
#C004B1.qcF <- C004B1[, C004B1$nFeature_Spatial > 60]
#C006C1.qcF <- C006C1[, C006C1$nFeature_Spatial > 100]
#C006D1.qcF <- C006D1[, C006D1$nFeature_Spatial > 100]

#Z <- c(A004A1.qcF, A004B1.qcF, A006C1.qcF, A006D1.qcF, C004A1.qcF, C004B1.qcF, C006C1.qcF, C006D1.qcF)

#Put Z into lapply below when testing different thresholds and X to visualize the general quality before quality control (QC).
grid.arrange(grobs=lapply(X, function(x) {
  SpatialFeaturePlot(x, features = "nFeature_Spatial") + 
    guides(fill= guide_colorbar(barwidth = 10,
                                title.position = "top",
                                label.vjust = 1.5,
                                ))}), ncol=8)

X <- makeX()
#Remove hemoglobin genes
clean <- lapply(X, function(x) {
  x <- x[!grepl("^HB.*-", rownames(x)), ]
  return(x)
})

list2env(clean, envir = globalenv())
rm(clean, Z)

```

```{r QC-Violin-and-spatial-feature-plots, results='hide', fig.dpi=300, fig.width=20, fig.height=6}
A004A1 <- A004A1[, A004A1$nFeature_Spatial > 40 & A004A1$percent_mito < 15 & A004A1$percent_ribo < 20] 
A004B1 <- A004B1[, A004B1$nFeature_Spatial > 40 & A004B1$percent_mito < 15 & A004B1$percent_ribo < 20] 
A006C1 <- A006C1[, A006C1$nFeature_Spatial > 90 & A006C1$percent_mito < 15 & A006C1$percent_ribo < 20] 
A006D1 <- A006D1[, A006D1$nFeature_Spatial > 90 & A006D1$percent_mito < 15 & A006D1$percent_ribo < 20] 
C004A1 <- C004A1[, C004A1$nFeature_Spatial > 60 & C004A1$percent_mito < 15 & C004A1$percent_ribo < 20] 
C004B1 <- C004B1[, C004B1$nFeature_Spatial > 60 & C004B1$percent_mito < 15 & C004B1$percent_ribo < 20] 
C006C1 <- C006C1[, C006C1$nFeature_Spatial > 100 & C006C1$percent_mito < 15 & C006C1$percent_ribo < 20] 
C006D1 <- C006D1[, C006D1$nFeature_Spatial > 100 & C006D1$percent_mito < 15 & C006D1$percent_ribo < 20] 


X <- makeX()

lapply(X, function(x) { #Violinplots
  VlnPlot(x, features = c("nCount_Spatial", "nFeature_Spatial", "percent_mito", "percent_ribo"), pt.size = 0.1, ncol = 4) + 
    plot_annotation(title = "Filtered reads for sample", subtitle = getObjName(x))
})


lapply(X, function(x) { #Spatial feature plots
  SpatialFeaturePlot(x, features = c("nCount_Spatial", "nFeature_Spatial", "percent_mito", "percent_ribo"), ncol = 4, pt.size.factor = 2.5) + 
    plot_annotation(title = "Filtered spots", subtitle = getObjName(x)) &
    guides(fill= guide_colorbar(barwidth = 10,
                                title.position = "top",
                                label.vjust = 1.5))
})

```

Normalization

```{r Normalization, results='hide'}

normalized_data <- lapply(X, function(x) {
  x <- NormalizeData(x, assay = "Spatial", verbose = T)
  x <- FindVariableFeatures(x)
  x <- ScaleData(x, features = rownames(x))
  return(x)
})

Y <- c("A004A1qc", "A004B1qc", "A006C1qc", "A006D1qc", "C004A1qc", "C004B1qc", "C006C1qc", "C006D1qc") 
names(normalized_data) <- paste0(Y)
list2env(normalized_data, envir = globalenv())
rm(normalized_data)

makeXqc <- function(x){
  X <- c(A004A1qc, A004B1qc, A006C1qc, A006D1qc, C004A1qc, C004B1qc, C006C1qc, C006D1qc)
  names(X) <- paste0(Y)
  return(X)
}

X <- makeXqc()

```

Feature plots of genes of interest.

```{r Genes-of-interest, results='hide', fig.dpi=300, fig.width=12, fig.height=15}
A004A1_fp <- SpatialFeaturePlot(A004A1qc, features = c("CD74", "PI16", "APOE", "LY6E", "MGP", "C1QC", "COMP", "JCHAIN", "CXCL2", "HLA-DQA1", "HLA-DRB1", "ZFP36"), alpha = c(0.3, 1), pt.size.factor = 2.5) + 
  plot_annotation("GOI's for A004A1")
A006C1_fp <- SpatialFeaturePlot(A006C1qc, features = c("CD74", "PI16", "APOE", "LY6E", "MGP", "C1QC", "COMP", "JCHAIN", "CXCL2", "HLA-DQA1", "HLA-DRB1", "ZFP36"), alpha = c(0.3, 1), pt.size.factor = 2.5) + 
  plot_annotation("Genes of interest in Aorta")

A004B1_fp <- SpatialFeaturePlot(A004B1qc,features = c("CD74", "PI16", "APOE", "LY6E", "MGP", "C1QC", "COMP", "JCHAIN", "CXCL2", "HLA-DQA1", "HLA-DRB1", "ZFP36"), alpha = c(0.3, 1), pt.size.factor = 2.5) +
  plot_annotation("GOI's for A004B1")
A006D1_fp <- SpatialFeaturePlot(A006D1qc,features = c("CD74", "PI16", "APOE", "LY6E", "MGP", "C1QC", "COMP", "JCHAIN", "CXCL2", "HLA-DQA1", "HLA-DRB1", "ZFP36"), alpha = c(0.3, 1), pt.size.factor = 2.5) +
  plot_annotation("GOI's for A006D1")

C004A1_fp <- SpatialFeaturePlot(C004A1qc, features = c("CD74", "PI16", "APOE", "LY6E", "MGP", "C1QC", "COMP", "JCHAIN", "CXCL2", "HLA-DQA1", "HLA-DRB1", "ZFP36"), alpha = c(0.3, 1), pt.size.factor = 2.5) + 
  plot_annotation("GOI's for C004A1")
C006C1_fp <- SpatialFeaturePlot(C006C1qc, features = c("CD74", "PI16", "APOE", "LY6E", "MGP", "C1QC", "COMP", "JCHAIN", "CXCL2", "HLA-DQA1", "HLA-DRB1", "ZFP36"), alpha = c(0.3, 1), pt.size.factor = 2.5) + 
  plot_annotation("GOI's for C006C1")

C004B1_fp <- SpatialFeaturePlot(C004B1qc,features = c("CD74", "PI16", "APOE", "LY6E", "MGP", "C1QC", "COMP", "JCHAIN", "CXCL2", "HLA-DQA1", "HLA-DRB1", "ZFP36"), alpha = c(0.3, 1), pt.size.factor = 2.5) +
  plot_annotation("Genes of interest in Carotid artery")
C006D1_fp <- SpatialFeaturePlot(C006D1qc,features = c("CD74", "PI16", "APOE", "LY6E", "MGP", "C1QC", "COMP", "JCHAIN", "CXCL2", "HLA-DQA1", "HLA-DRB1", "ZFP36"), alpha = c(0.3, 1), pt.size.factor = 2.5) +
  plot_annotation("GOI's for C006D1")

A004B1_fp  
A006C1_fp
A004A1_fp  
A006D1_fp
C004B1_fp  
C006C1_fp
C004A1_fp  
C006D1_fp

ggsave(plot = A006C1_fp, filename = "GOIs_aorta.png", device = "png", width = 30, height = 30, units = "cm", dpi = 300)
ggsave(plot = C004B1_fp, filename = "GOIs_ca.png", device = "png", width = 30, height = 30, units = "cm", dpi = 300)

```

#Run PCA
```{r Principal-Component-Analysis, results='hide', fig.dpi=300, fig.width=8, fig.height=7}

addpca <- lapply(X, function(x) {
  x <- RunPCA(x, verbose = FALSE)
  return(x)
})

lapply(addpca, ElbowPlot)

names(addpca) <- paste0(Y)
list2env(addpca, envir = globalenv())
rm(addpca)

X <- makeXqc()

```

Find neighbors and clusters and run UMAP.
```{r Find-neighbors-clusters-and-run-UMAP, results='hide'}

ncu_all <- lapply(X, function(x) {
  x <- FindNeighbors(x, reduction = "pca", dims = 1:15) #The number of dimension was decided based on elbow plots. There weren't huge differences so an average across the samples is used for the maximum.
  x <- RunUMAP(x, reduction = "pca", dims = 1:15)
  return(x)
})

names(ncu_all) <- paste0(Y)
list2env(ncu_all, envir = globalenv())
rm(ncu_all)

X <- makeXqc()


```


Run Clustree to determine the appropriate number of clusters 
```{r Cluster-tree, results='hide', fig.dpi=300, fig.width=8, fig.height=9}
clusters <- lapply(X, function(x){ #This run FindClusters with a range of resolutions that are saved into metadata and can be then used with the clustree function.
  resolution.range <- seq(from = 0, to = 1, by = 0.1)
  x <- FindClusters(x, resolution = resolution.range)
})


names(clusters) <- paste0(Y)
list2env(clusters, envir = globalenv())

lapply(clusters, function(x) {
  clustree(x) + plot_annotation(title = " Clustertree for samples", subtitle = getObjName(x))
} ) 

rm(clusters)
X <- makeXqc()
```


Plot UMAP and spatial dimensional plot (clustering on tissue)

```{r Clusters, results='hide', fig.dpi=300, fig.width=30, fig.height=15}

X <- lapply(X , function(x) {
  Idents(x) <- "Spatial_snn_res.0.8"
  return(x)
  })

names(X) <- paste0(Y)
list2env(X, envir = globalenv())

```

Differential gene expression
```{r Differential-Gene-Expression, results='hide', fig.dpi=300, fig.width=20, fig.height=15}

Y <- c("A004A1_markers", "A004B1_markers", "A006C1_markers", "A006D1_markers", "C004A1_markers", "C004B1_markers", "C006C1_markers", "C006D1_markers")

Markers <- lapply(X, function(x) {
  Idents(x) <- "Spatial_snn_res.0.8"
  FindAllMarkers(x)
})
names(Markers) <- paste0(Y)
list2env(Markers, envir = globalenv())
rm(Markers)

#Save the markers lists to CSVs.
for(i in 1:length(Y)) {                
  write.csv(get(Y[i]),
             paste("~/Desktop/TERVA/TERVA2DATA/SPATIAL 2022/",
                   Y[i],
                   ".csv"),
             row.names = TRUE)
}

#Pick top logFC genes and make heatmap
#Rerun GSEA with new marker lists

top10 <- A004A1_markers %>%
  subset(p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  top_n(n = 10, wt = avg_log2FC)


clusterheatmap <- DoHeatmap(subset(A004A1qc, downsample=500), features = top10$gene, size = 4, angle = 90, label = T, slot = "scale.data") + NoLegend() + 
  theme(axis.text.y = element_text(size=12)) 

p1 <- SpatialFeaturePlot(object = A004A1qc, features = rownames(A004A1_markers)[1:9], alpha = c(0.3, 1)) +
  plot_annotation(title = "DEGs for A004A1")
p2 <- SpatialFeaturePlot(object = A004B1qc, features = rownames(A004B1_markers)[1:9], alpha = c(0.3, 1)) +
  plot_annotation(title = "DEGs for A004B1")  
p3 <- SpatialFeaturePlot(object = A006C1qc, features = rownames(A006C1_markers)[1:9], alpha = c(0.3, 1)) +
  plot_annotation(title = "DEGs for A006C1")
p4 <- SpatialFeaturePlot(object = A006D1qc, features = rownames(A006D1_markers)[1:9], alpha = c(0.3, 1)) +
  plot_annotation(title = "DEGs for A006D1")
p5 <- SpatialFeaturePlot(object = C004A1qc, features = rownames(C004A1_markers)[1:9], alpha = c(0.3, 1)) +
  plot_annotation(title = "DEGs for C004A1")
p6 <- SpatialFeaturePlot(object = C004B1qc, features = rownames(C004B1_markers)[1:9], alpha = c(0.3, 1)) +
  plot_annotation(title = "DEGs for C004B1")  
p7 <- SpatialFeaturePlot(object = C006C1qc, features = rownames(C006C1_markers)[1:9], alpha = c(0.3, 1)) +
  plot_annotation(title = "DEGs for C006C1")
p8 <- SpatialFeaturePlot(object = C006D1qc, features = rownames(C006D1_markers)[1:9], alpha = c(0.3, 1)) +
  plot_annotation(title = "DEGs for C006D1")

p1
p3 
p2 
p4
p5 
p7 
p6 
p8
```

Compare aorta and carotid artery

```{r Tissue comparison, results='hide'}

combined_data <- merge(A004A1qc, c(A004B1qc, A006C1qc, A006D1qc, C004A1qc, C004B1qc, C006C1qc, C006D1qc))
combined_data <- JoinLayers(combined_data)

coldata <- as.data.frame(combined_data@meta.data$sample_id)
coldata <- mutate(coldata,
                  Tissue = case_when(
                    startsWith(combined_data@meta.data$sample_id, "AO") ~ "Aorta",
                    startsWith(combined_data@meta.data$sample_id, "CA") ~ "Carotid_artery"
                  ))
tissue_id <- coldata$Tissue
combined_data@meta.data$tissue_id <- as.factor(tissue_id)

Idents(combined_data) <- "tissue_id"
combined_data <- ScaleData(combined_data, features = rownames(combined_data))

AOvsCA_markers <- FindMarkers(combined_data, only.pos = F, ident.1 = "Aorta", ident.2 = "Carotid_artery", min.pct = 0.05)

only_sig <- AOvsCA_markers %>% subset(p_val_adj < 0.05 & avg_log2FC > 1 | avg_log2FC < -1)
write.csv(only_sig, file = "AOvsCAmarkers.csv", sep = ";", quote = F, row.names = T, dec = ",")


```
```{r volcanoplot, fig.dpi=300, fig.width=15, fig.height=15}
AO_CA_volcano <- EnhancedVolcano::EnhancedVolcano(AOvsCA_markers,
                          lab = rownames(AOvsCA_markers),
                          x = 'avg_log2FC',
                          y = 'p_val_adj',
                          title = 'Differentially expressed genes between human Aorta and Carotid artery',
                          pCutoff = 1e-2,
                          FCcutoff = 0.4,
                          pointSize = 3.0,
                          col=c('grey', 'slateblue2', 'cyan2', 'cyan4'),
                          labSize = 6.0,
                          titleLabSize = 15.0)

ggsave(plot = AO_CA_volcano, filename = "AorstavsCAvolcanoplot.png", device = "png", path = getwd(), dpi = 300, width = 30, height = 30, units = "cm")

```

```{r Immunoglobulins, results='hide', fig.dpi=300, fig.width=12, fig.height=15}

A006C1_IG <- SpatialFeaturePlot(A006C1qc, features = c("IGHG2", "IGHG4", "IGLC7","IGLC1", "IGKC", "IGHG3"), alpha = c(0.3, 1), pt.size.factor = 2.5) + 
  plot_annotation("Immunoglobulin expression in Aorta")



C004A1_IG <- SpatialFeaturePlot(C004A1qc,features = c("IGHG2", "IGHG4", "IGLC7","IGLC1", "IGKC", "IGHG3"), alpha = c(0.3, 1), pt.size.factor = 2.5) +
  plot_annotation("Immunoglobulin expression in Carotid artery")


 
A006C1_IG
C004A1_IG  

ggsave(plot = A006C1_IG, filename = "Immunoglobulinsinaorta.png", device = "png", path = getwd(), dpi = 300, width = 20, height = 20, units = "cm")

ggsave(plot = C004A1_IG, filename = "Immunoglobulinsinca.png", device = "png", path = getwd(), dpi = 300, width = 20, height = 20, units = "cm")
```

```{r CD74_macrophages, results='hide', fig.dpi=300, fig.width=12, fig.height=15}

A006C1_CD74 <- SpatialFeaturePlot(A006C1qc, features = c("CD74", "CD68"), alpha = c(0.3, 1), pt.size.factor = 2.5) + 
  plot_annotation("CD74 and macrophage marker expression in Aorta")



C004A1_CD74 <- SpatialFeaturePlot(C004A1qc,features = c("CD74", "CD68"), alpha = c(0.3, 1), pt.size.factor = 2.5) +
  plot_annotation("CD74 and macrophage marker expression in Carotid artery")


 
A006C1_CD74
C004A1_CD74  

comb_cd74 <- A006C1_CD74 / C004A1_CD74 + plot_annotation(title = "CD74 and CD68 expression in Aorta (Top) and Carotid artery (bottom)")

ggsave(plot = comb_cd74, filename = "CD74inaortaandca.png", device = "png", path = getwd(), dpi = 300, width = 20, height = 20, units = "cm")


```

```{r complement-genes, results='hide'}

A006C1_CC <- SpatialFeaturePlot(A006C1qc, features = c("C4B", "C3", "C1QA", "C1QB", "C1QC","CFH","C1S", "C7"), alpha = c(0.3, 1), pt.size.factor = 2.5) + 
  plot_annotation("Complement cascade component expression in Aorta")



C004A1_CC <- SpatialFeaturePlot(C004A1qc,features = c("C4B", "C3", "C1QA", "C1QB", "C1QC","CFH","C1S", "C7"), alpha = c(0.3, 1), pt.size.factor = 2.5) +
  plot_annotation("Complement cascade component expression in  Carotid artery")


 
A006C1_CC
C004A1_CC  

comb_cc <- (A006C1_CC) | (C004A1_CC) + plot_annotation(title = "Complement cascade component expression in Aorta (left) and Carotid artery (right)")

ggsave(plot = comb_cc, filename = "Complementactivation_aorta_ca.png", device = "png", path = getwd(), dpi = 300, width = 30, height = 20, units = "cm")


```