---
title: "TERVA SPATIAL CAROTID ARTERY 2022"
author: "Lea Mikkola"
date: "23/06/2022"
output:
  html_document:
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

setwd("/scratch/project_2005050/Rstats/Spatial/")

.libPaths(c("/projappl/project_2005050/project_rpackages_4.1.1", .libPaths()))

library(Matrix)
library(dplyr)
library(Seurat)
library(ggplot2)
library(patchwork)
library(hdf5r)
library(viridis)
library(clustree)

```

First load the data in with the Load_spatial R script.

*QC*

```{r}
X <- c(C004A1, C004B1, C006C1, C006D1)

L <- lapply(X, function(x) {
  x@meta.data[["percent_mito"]] <- PercentageFeatureSet(x, "^MT-")
  x@meta.data[["percent_ribo"]] <- PercentageFeatureSet(x, "^RP[SL]")
  return(x)
})

Y <- c("C004A1", "C004B1", "C006C1", "C006D1") 
names(L) <- paste0(Y)
list2env(L, envir = globalenv())
rm(L)

C004A1.qcF <- C004A1[, C004A1$nFeature_Spatial > 50] #Test different thresholds and visually inspect what would be suitable level.
C004B1.qcF <- C004B1[, C004B1$nFeature_Spatial > 50]
C006C1.qcF <- C006C1[, C006C1$nFeature_Spatial > 90]
C006D1.qcF <- C006D1[, C006D1$nFeature_Spatial > 100]

X <- c(C004A1.qcF, C004B1.qcF, C006C1.qcF, C006D1.qcF)
do.call(gridExtra::grid.arrange, c(lapply(X, SpatialFeaturePlot, features = "nFeature_Spatial"), ncol=4))
```

```{r}
C004A1 <- C004A1[, C004A1$nFeature_Spatial > 50 & C004A1$percent_mito < 15 & C004A1$percent_ribo < 20] 
C004B1 <- C004B1[, C004B1$nFeature_Spatial > 50 & C004B1$percent_mito < 15 & C004B1$percent_ribo < 20] 
C006C1 <- C006C1[, C006C1$nFeature_Spatial > 90 & C006C1$percent_mito < 15 & C006C1$percent_ribo < 20] 
C006D1 <- C006D1[, C006D1$nFeature_Spatial > 100 & C006D1$percent_mito < 15 & C006D1$percent_ribo < 20] 

X <- c(C004A1, C004B1, C006C1, C006D1)

lapply(X, function(x) { #Violinplots
  VlnPlot(x, features = c("nCount_Spatial", "nFeature_Spatial", "percent_mito", "percent_ribo"), pt.size = 0.1, ncol = 4) + 
    NoLegend() + plot_annotation(title = "Filtered reads")
})

lapply(X, function(x) { #Spatial feature plots
  SpatialFeaturePlot(x, features = c("nCount_Spatial", "nFeature_Spatial", "percent_mito", "percent_ribo"), ncol = 4) + 
    NoLegend() + plot_annotation(title = "Filtered spots")
})

```

Identify the highly expressed genes/marker genes.

```{r}
Identify_markers <- function(x) {
C = x@assays$Spatial@counts 
C@x = C@x/rep.int(colSums(C), diff(C@p)) # calculate the sum of the count for each gene and compare them together 
most_expressed <- order(Matrix::rowSums(C), decreasing = T)[20:1] # Put the top 20 genes in order /high to low counts 
boxplot(as.matrix(t(C[most_expressed, ])), cex = 0.1, las = 1, xlab = "% total count per cell", 
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE)
}

lapply(X, Identify_markers)

```

Normalization

```{r}

normalized_data <- lapply(X, function(x) {
  x <- SCTransform(x, assay = "Spatial", verbose = T)
  return(x)
})

Y <- c("C004A1qc", "C004B1qc", "C006C1qc", "C006D1qc") 
names(normalized_data) <- paste0(Y)
list2env(normalized_data, envir = globalenv())
rm(normalized_data)

X <- c(C004A1qc, C004B1qc, C006C1qc, C006D1qc)


```

Feature plots of genes of interest

```{r}
C004A1_fp <- SpatialFeaturePlot(C004A1qc, slot = "data", features = c("LGALS3", "MGP", "CD74", "TIMP1", "SPP1", "FTL"), alpha = c(0.3, 1)) + 
  plot_annotation("GOI's for C004A1")
C006C1_fp <- SpatialFeaturePlot(C006C1qc, slot = "data", features = c("LGALS3", "MGP", "CD74", "TIMP1", "SPP1", "FTL"), alpha = c(0.3, 1)) + 
  plot_annotation("GOI's for C006C1")

C004B1_fp <- SpatialFeaturePlot(C004B1qc, slot = "data",features = c("LGALS3", "MGP", "CD74", "TIMP1", "SPP1", "FTL"), alpha = c(0.3, 1)) +
  plot_annotation("GOI's for C004B1")
C006D1_fp <- SpatialFeaturePlot(C006D1qc, slot = "data",features = c("LGALS3", "MGP", "CD74", "TIMP1", "SPP1", "FTL"), alpha = c(0.3, 1)) +
  plot_annotation("GOI's for C006D1")

C004A1_fp  
C006C1_fp
C004B1_fp  
C006D1_fp
```

```{r}
C004B1_fp <- SpatialFeaturePlot(C004B1qc, slot = "data", features = c("CD74", "SPP1"), alpha = c(0.3, 1)) + 
  plot_annotation("CD74 and SPP1 expression in sample B")
```


```{r}
Lgals3_interact <- SpatialFeaturePlot(C004A1, slot = "data", features = c("LGALS3"), interactive = TRUE)
l1 <- SpatialFeaturePlot(C004B1qc, features = c("LGALS3"), alpha = c(0.6,1))
l2 <- SpatialFeaturePlot(C006D1qc, features = c("LGALS3"), alpha = c(0.6,1), pt.size.factor = 1.9)
l3 <- SpatialFeaturePlot(C004B1qc, features = c("LGALS3"), alpha = c(0,0))
l4 <- SpatialFeaturePlot(C006D1qc, features = c("LGALS3"), alpha = c(0,0))

l1 + l3 
l2 + l4
```

Feature plots of macrophage markers
```{r}

C004A1_mcphg <- SpatialFeaturePlot(C004A1qc, features = c("CD68", "CD163", "CCL2", "MCAM", "MRC1", "LGALS3"), alpha = c(0.3, 1)) + 
  plot_annotation("Macrophage's in C004A1")
C006C1_mcphg <- SpatialFeaturePlot(C006C1qc, features = c("CD68", "CD163", "CCL2", "MCAM", "MRC1", "LGALS3"), alpha = c(0.3, 1)) + 
  plot_annotation("Macrophage's in C006C1")

C004B1_mcphg <- SpatialFeaturePlot(C004B1qc, features = c("CD68", "CD163", "CCL2", "MCAM", "MRC1", "LGALS3"), alpha = c(0.3, 1)) +
  plot_annotation("Macrophage's in C004B1")
C006D1_mcphg <- SpatialFeaturePlot(C006D1qc, features = c("CD68", "CD163", "CCL2", "MCAM", "MRC1", "LGALS3"), alpha = c(0.3, 1)) +
  plot_annotation("Macrophage's in C006D1")

C004A1_mcphg
C006C1_mcphg
C004B1_mcphg  
C006D1_mcphg

```

#Run PCA
```{r}

addpca <- lapply(X, function(x) {
  x <- RunPCA(x, assay = "SCT", verbose = FALSE)
  return(x)
})

lapply(addpca, ElbowPlot)

Y <- c("C004A1qc", "C004B1qc", "C006C1qc", "C006D1qc")
names(addpca) <- paste0(Y)
list2env(addpca, envir = globalenv())
rm(addpca)

X <- c(C004A1qc, C004B1qc, C006C1qc, C006D1qc)



```

Find neighbors and clusters and run UMAP.
```{r}

ncu_all <- lapply(X, function(x) {
  x <- FindNeighbors(x, reduction = "pca", dims = 1:15) #The number of dimension was decided based on elbow plots. There weren't huge differences so an average across the samples is used for the maximum.
  x <- RunUMAP(x, reduction = "pca", dims = 1:15)
  return(x)
})

names(ncu_all) <- paste0(Y)
list2env(ncu_all, envir = globalenv())
rm(ncu_all)

X <- c(C004A1qc, C004B1qc, C006C1qc, C006D1qc)


```

Run Clustree to determine the appropriate number of clusters 
```{r}
clusters <- lapply(X, function(x){ #This run FindClusters with a range of resolutions that are saved into metadata and can be then used with the clustree function.
  resolution.range <- seq(from = 0, to = 1, by = 0.1)
  x <- FindClusters(x, resolution = resolution.range)
return(x)
})

names(clusters) <- paste0(Y)
list2env(clusters, envir = globalenv())


lapply(clusters, function(x) {
  clustree(x) + ggtitle(names(x), "Cluster tree")
} ) 
rm(clusters)

X <- c(C004A1qc, C004B1qc, C006C1qc, C006D1qc)

#FIX NAMING OF THE PLOTS LATER
```
Resolution 0.3 should be a good compromice across samples, but fix later if looks inappropriate for some samples.

Plot UMAP and spatial dimensional plot (clustering on tissue)

```{r}

do.call(gridExtra::grid.arrange, c(lapply(X, DimPlot, group.by = "SCT_snn_res.0.3", reduction = "umap", pt.size = 0.1), ncol=2))
do.call(gridExtra::grid.arrange, c(lapply(X, SpatialDimPlot, label = F, label.size = 1), ncol=4))


```

1. Identification of all markers for each cluster: this analysis compares each cluster against all others and outputs the genes that are differentially expressed/present
2. dentification of conserved markers for each cluster: This analysis looks for genes that are differentially expressed/present within each condition first, and then reports those genes that are conserved in the cluster across all conditions. These genes can help to figure out the identity for the cluster.
 

3. Identification of Variable Features:

```{r}
# differential expression genes in all clusters 

lapply(X, function(x) { #Only one SCT model found, recorrection skipped.
  x <- PrepSCTFindMarkers(x)
return(x)
})

Y <- c("C004A1_markers", "C004B1_markers", "C006C1_markers", "C006D1_markers")

Markers <- lapply(X, function(x) {
  FindAllMarkers(x)
})
names(Markers) <- paste0(Y)
list2env(Markers, envir = globalenv())
rm(Markers)
saveRDS(C004A1_markers, "C004A1_markers.rds")
saveRDS(C004B1_markers, "C004B1_markers.rds")
saveRDS(C006C1_markers, "C006C1_markers.rds")
saveRDS(C006D1_markers, "C006D1_markers.rds")

p1 <- SpatialFeaturePlot(object = C004A1qc, features = rownames(C004A1_markers)[1:12], alpha = c(0.3, 1)) +
  plot_annotation(title = "DEGs for C004A1")
p2 <- SpatialFeaturePlot(object = C004B1qc, features = rownames(C004B1_markers)[1:12], alpha = c(0.3, 1)) +
  plot_annotation(title = "DEGs for C004B1")  
p3 <- SpatialFeaturePlot(object = C006C1qc, features = rownames(C006C1_markers)[1:12], alpha = c(0.3, 1)) +
  plot_annotation(title = "DEGs for C006C1")
p4 <- SpatialFeaturePlot(object = C006D1qc, features = rownames(C006D1_markers)[1:12], alpha = c(0.3, 1)) +
  plot_annotation(title = "DEGs for C006D1")

p1
p3
p2
p4

```

Find Spatially variable features.

```{r}

SVF <- lapply(X, function(x) {
  x <- FindSpatiallyVariableFeatures(x, assay = 'SCT', selection.method = 'markvariogram',  features = VariableFeatures(x)[1:1000])
return(x)
})

Y <- c("C004A1_SVGs", "C004B1_SVGs", "C006C1_SVGs", "C006D1_SVGs") 
names(SVF) <- paste0(Y)
list2env(SVF, envir = globalenv())
rm(SVF)

X <- c(C004A1_SVGs, C004B1_SVGs, C006C1_SVGs, C006D1_SVGs) 
top.features <- lapply(X, function(x) {
  x <- head(SpatiallyVariableFeatures(x, selection.method = "markvariogram"), 20)
  return(x)
})

Y <- c("C004A1_SVGs", "C004B1_SVGs", "C006C1_SVGs", "C006D1_SVGs") 
names(top.features) <- paste0(Y)
list2env(top.features, envir = globalenv())
rm(top.features)


C004A1_SFP <- SpatialFeaturePlot(C004A1qc, features = C004A1_SVGs[1:10], ncol = 5, alpha = c(0.3, 1)) + 
  plot_annotation(title = "Spatially variable features C004A1")
  
C004B1_SFP <- SpatialFeaturePlot(C004B1qc, slot = "data", features = C004B1_SVGs[1:10], ncol = 5, alpha = c(0.3, 1)) +
  plot_annotation(title = "Spatially variable features C004B1")

C006C1_SFP <- SpatialFeaturePlot(C006C1qc, features = C006C1_SVGs[1:10], ncol = 5, alpha = c(0.3, 1)) +
  plot_annotation(title = "Spatially variable features C006C1")

C006D1_SFP <- SpatialFeaturePlot(C006D1qc, slot = "data", features = C006D1_SVGs[1:10], ncol = 5, alpha = c(0.3, 1)) +
  plot_annotation(title = "Spatially variable features C006D1")

C004B1_SFP  
C006D1_SFP

```
